Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-08-09 manuscript/coverletter/collabra_response.tex
work on response letter

		This reviewer seems to strongly distinguish between students and investigators. Perhaps this is from a misconception of who we are talking about when we talk about students (perhaps due to differences in training in different areas). To clarify, these were replications done primarily by first year PhD students at an elite US institution. Many publications (including in top journals) are in fact done by PhD students, under the supervision of professors. We checked the google scholars of the rescuers here and 7/17 of them (40\%) already had peer-reviewed journal publications, in some cases multiple. These students are or will soon be publishing in the literature. We agree that they may not have extensive expertise, but we think they and thus their rescue projects are reasonably representative of work by early career researchers with limited resources.

@@

		The original replications were performed under limitations, such as time limits and online samples, which we acknowledge, but which we believe is common when early career researchers build on existing work.

		As for when original materials aren't available, when they aren't available, the replications are as close as they can be, which may be not very close. Replicators did what they could given the external constrain of unavailable materials.

		[TODO not really sure what to say here]

@@

		[TODO not sure what to do here]

@@

		There are many metrics for what counts as a successful replication, but many of them depend on a key-statistic. As discussed in Hawkins et al 2018 (https://journals.sagepub.com/doi/full/10.1177/2515245917740427) and Boyce et al 2023 (https://royalsocietypublishing.org/doi/full/10.1098/rsos.231240), original papers often report a number of effects, and it is not always clear which one is the most key. Additionally, often the pattern of interest is dependent on multiple numeric effects (ex. two main effects and their interaction). Thus we think the holistic, subjective assessment of whether or not a study replicated is important, although we also do analyze the key effect as well.

@@

		As we write the in the methods section, ``Each project was rated on the basis of subjective replication success (on a 0--1 scale) by both MCF and one of VB and BP.
		We thought about replication in terms of general match in key study results (direction, magnitude, and significance), rather than focusing on any singular numeric result or significance cut-off.''

@@

		Whether a replication is statistically significant in the same direction as the original is one metric that is sometimes used as a measure of replicability, but as TODO CITE SOME PEOPLE SAYING THIS IS DUMB. Patil 2016, Simonsohn 2015, ANdrews and Kasy 2019 are potential cites here!

@@

		The effect sizes shown are SMDs, but in order to show everything on the same scale, we had to force everything onto that scale, even if it required a lot of conversions requiring approximations. For assessing individual studies it makes more sense to look at their raw results, rather the converted results.

@@

		We are fairly confident that there is not fraud as a) committing fraud would have been strictly more work than not and b) students had no incentive to commit fraud. During the class, students submitted their projects for a number of checkpoints. Thus, we know that students had working implementations of the studies because the TAs went through every experiment and gave feedback. Additionally, students had to have working code that ran on their pilot data before they were approved for data collection. We also know that the students ran studies on the class Prolific account because we can see how many participants were recruited and paid for each study. Thus, we know data was collected. Students knew that their grades in no way depended on what results they got (i.e. whether the study replicated), so had no incentive to change the data they had collected.

@@

		Thanks for this suggestion. In our added case study expansion, we now mention the additional studies and where they come from, citing those from the published literature.

		%TODO this ^

@@

		I think there may be a misunderstanding here -- we aren't doing meta-analyses because we aren't interested in the aggregate effect size across the original and replications.

		When we discuss the "meta-analytic" mean of some studies that is because the measures of consistency we use (p-original) is between two things, so to compare more than two, some have to be pooled (which does make use of meta-analysis to do the pooling to get an aggregate mean and spread). As described in Mathur and VanderWeele 2020 (https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssa.12572) which introduces the statistic of p-original, it uses the point estimate and standard error of the original compared to the average population effect size in the replications and standard error thereof. As they mention, the most common way to get the average population effect size is via a random effects meta analysis.

@@

		As we are only interested in studies that are direct replications of the original, we think we got them. 
		%TODO how to navigate this s

		%TODO add
		We understand the concern about wanting to see just the replication and rescue, so we have added a column to TABLE TODO with this.

@@

		 Again, it seems like there may be confusion here -- the "p-values" presented here are "p-original" which is p-value like in that it's a measure of how likely an estimate would be to be at least as extreme as observed given a "null" distribution. It is a measure of statistical (in-)consistency on the effect sizes, not a measure of how close to zero the effect sizes are.

@@

		We apologize for this oversight and have added to the caption short descriptions for these two as well as a cross reference to where the predictors are discussed at length in Boyce et al 2023.

		%TODO DO THIS 
		\theysaid{It is the open science collaborations, not consortium.}

		%TODO FIX
		Oops, thank you for bringing this to our attention, we have fixed it.

@@

		This point is well taken and we have expanded our discussion to acknowledge the larger space of factors and decisions at play with replications and rescues and the motivations to do them. 
		%TODO make not suck

@@

		%TODO FIX
		We have clarified this. 
		\theysaid{So in general, I think this is a neat little paper, which, in my opinion, could benefit from adding a little bit more discussion/depth in the end on what its outcomes could entail.}

@@

		We again would like to thank both the editor and the reviewers for their thoughtful and prompt reviews. We hope that our changes to the manuscript and the documentation of our repository offer the necessary clarifications and nuance.

>>>>

		This reviewer seems to strongly distinguish between students and investigators. Perhaps this is from a misconception of who we are talking about when we talk about students (perhaps due to differences in training in different areas).

		To clarify, these were replications done primarily by first year PhD students at an elite US institution. Many publications (including in top journals) are in fact done by PhD students, under the supervision of professors. We checked the google scholars of the rescuers here and 7/17 of them (40\%) already had peer-reviewed journal publications, several of which were first-authored. These students are or will soon be publishing in the literature. We agree that they may not have extensive expertise, but we think they and thus their rescue projects are reasonably representative of work by early career researchers with limited resources.

		Since ``student'' can bring to mind different populations for different people, we have clarified in our limitations section who the rescuers were and how we think this is likely to generalize to other early career researchers (i.e. many of the people conducting research) but may not generalize to experts.

		The limitations section reads:

		\revised{
			The reported rescue projects are a small sample of replications, leading to highly uncertain estimates.
			Additionally, they were chosen non-randomly, as they were doubly selected for based on student interest -- once for the original replication and again for the rescue.
			However, this selection bias is likely to correlate with how graduate students and other early career researchers choose what topics to work on and what studies to build on.
			Nevertheless, our estimate is still broadly consistent with that of ManyLabs 5 (29\% vs.~22\%, Ebersole et al., 2020).
			\newline

			One potentially salient limitation of our work here is that the rescue projects (and the first replications) were primarily conducted by first year PhD students; hence the generality of our conclusions to other investigators is unknown. We believe, however, that this is more of an apparent than an actual limitation. Students were enrolled in a highly-competitive graduate program where it is expected that PhD students author peer-reviewed papers; several of the rescuers already had published first-authored peer reviewed articles. Thus, while we refer to the rescuers as students, they are investigators doing experiments that become part of the literature.
			\newline

			Students conducting rescue projects put substantial effort into trying to set up rescues that had a good chance of success, but projects were constrained by budget limitations, a short timeline, and the constraint that most data collection needed to occur online. Additionally, while most students had done research before and were learning best practices for experiments, they were not experts in the particular paradigms they were implementing. These limitations are representative of the sort of knowledge and resource limitations often faced by early-career researchers. That said, it is possible that different results might be obtained in better-resourced settings, by scientists with more expertise, more time, or more privileged knowledge about how the original experiments were conducted.
			\newline

			Due to the limited number of replications that are conducted and systematically reported, and the even smaller number of re-replications reported, there is substantial uncertainty in how much replicability varies across features of the original experiment and replications.}

@@

		The original replications were performed under limitations, such as time limits and online samples, which we acknowledge, but which we believe is common when early career researchers build on existing work. They may be limited in what resources (time, money, access to populations) are available, and so what results are accessible to them is relevant.

		As for when original materials aren't available, the replications are as close as they can be, which may be not very close. Replicators did what they could given the external constraint of unavailable materials.

@@

		The idea behind 2.5x sample size is that it gives the replication an 80\% chance of correctly rejecting the null hypothesis of an effect that the original was 33\% powered for. This standard has the convenience of being agnostic to the type of experiment and test being conducted (unlike calculating effect sizes or calculating power, which both can be difficult and very sensitive to design). This 2.5x is a target for answering is there an effect big enough for the original study to have reasonably detected.

@@

		There are many metrics for what counts as a successful replication, but many of them depend on a key-statistic. As discussed in Hawkins et al 2018 (https://journals.sagepub.com/doi/full/10.1177/2515245917740427) and Boyce et al 2023 (https://royalsocietypublishing.org/doi/full/10.1098/rsos.231240), original papers often report a number of effects, and it is not always clear which one is the most key. Additionally, often the pattern of interest is dependent on multiple numeric effects (ex. two main effects and their interaction). Thus we think the holistic, subjective assessment of whether or not a study replicated is important, although we also do analyze a key effect as well.

@@

		As we wrote the in the methods section, ``Each project was rated on the basis of subjective replication success (on a 0--1 scale) by both MCF and one of VB and BP.
		We thought about replication in terms of general match in key study results (direction, magnitude, and significance), rather than focusing on any singular numeric result or significance cut-off.''

@@

		Whether a replication is statistically significant in the same direction as the original is one metric that is sometimes used as a measure of replicability, but as noted by Patil et al 2016 (doi.org/10.1177/1745691616646366), it isn't a very good one as differences in where $p<.05$ can arise from sampling variation.

@@

		The effect sizes shown are SMDs, but in order to show the effects on the same scale, we had to coerce all the numbers onto the same scale, even if it required approximations as part of the conversions. For assessing individual studies it makes more sense to look at their raw results, rather the converted results.

@@

		We are fairly confident that there is not fraud as a) committing fraud would have been strictly more work than not and b) students had no incentive to commit fraud. During the class, students submitted their projects for a number of checkpoints. Thus, we know that students had working implementations of the studies because the TAs went through every experiment and gave feedback. Additionally, students had to have working code that ran on their pilot data before they were approved for running the main data collection. We also know that the students ran studies on the class Prolific account because we can see how many participants were recruited and paid for each study. Thus, we know data was collected. Students knew that their grades in no way depended on what results they got (i.e. whether the study replicated), so had no incentive to change the data they had collected.

@@

		Thanks for this suggestion. In the case study section and the appendix, when we describe studies where there were additional replications, we now mention the additional studies and where they come from, citing those from the published literature.

@@

		I think there may be a misunderstanding here -- we aren't doing meta-analyses in order to aggregate effect size across the the original and replications.

		When we discuss the ``meta-analytic'' mean of some studies that is because the measures of consistency we use (p-original) is between two things, so to compare more than two, some have to be pooled (which does make use of meta-analysis to do the pooling to get an aggregate mean and spread). As described in Mathur and VanderWeele 2020 (https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssa.12572) which introduces the statistic of p-original, it uses the point estimate and standard error of the original compared to the average population effect size in the replications and standard error thereof. As they mention, the most common way to get the average population effect size is via a random effects meta analysis.

@@

		Students did search for direct replications of the studies, which if they are in the published literature are likely to be found. It is possible that there happen to be direct replications that aren't labeled as such that we could have missed, and of course, there could be unpublished or grey literature replications.

		We understand the concern about wanting to see just the replication and rescue, so we have added a column to Table 1 with this information.

		New Table 1 shown below:

			\begin{tabular}[t]{lrrrrr}
				\toprule
				\multicolumn{1}{c}{ } & \multicolumn{5}{c}{p-original comparing between} \\
				\cmidrule(l{3pt}r{3pt}){2-6}
				\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Original and} & \multicolumn{1}{c}{Rescue and} \\
				\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-6}
				Paper & All reps & 1st Rep and Rescue & Rescue & Non-rescue & Other reps\\
				\midrule
				Birch \& Bloom 2007 & 0.192 & 0.189 & 0.194 & 0.191 & 0.989\\
				Child et al. 2018 & 0.669 & 0.669 & 0.641 & 0.858 & 0.778\\
				Chou et al. 2016 & 0.055 & 0.055 & 0.101 & 0.005 & 0.231\\
				Craig \& Richeson 2014 & 0.145 & 0.145 & 0.001 & 0.302 & 0.020\\
				Gong et al. 2019 & 0.262 & 0.262 & 0.057 & 0.511 & 0.228\\
				Haimovitz \& Dweck 2016 & 0.068 & 0.018 & 0.018 & 0.180 & 0.867\\
				Hopkins et al. 2016 & 0.290 & 0.290 & 0.654 & 0.004 & 0.015\\
				Jara-Ettinger et al. 2022 & 0.014 & 0.014 & 0.000 & 0.006 & 0.000\\
				Krauss \& Wang 2003 & 0.271 & 0.271 & 0.519 & 0.139 & 0.311\\
				Ngo et al. 2019 & 0.106 & 0.000 & 0.000 & 0.385 & 0.257\\
				Paxton et al. 2012 & 0.011 & 0.011 & 0.005 & 0.023 & 0.649\\
				Payne et al. 2008 & 0.023 & 0.071 & 0.031 & 0.079 & 0.895\\
				Porter et al. 2016 & 0.370 & 0.370 & 0.905 & 0.002 & 0.003\\
				Schechtman et al. 2010 & 0.712 & 0.712 & 0.654 & 0.770 & 0.877\\
				Tarampi et al. 2016 & 0.000 & 0.000 & 0.000 & 0.000 & 0.145\\
				Todd et al. 2016 & 0.062 & 0.100 & 0.124 & 0.058 & 0.778\\
				Yeshurun \& Levy 2003 & 0.614 & 0.007 & 0.017 & 0.943 & 0.474\\
				\bottomrule
			\end{tabular}

@@

		 Again, it seems like there may be confusion here -- the ``p-values'' presented here are ``p-original'' which is p-value like in that it's a measure of how likely an estimate would be to be at least as extreme as observed given a ``null'' distribution. It is a measure of statistical (in-)consistency on the effect sizes, not a measure of how close to zero the effect sizes are.

@@

		We apologize for this oversight. We have changed these row names to ``Original authors at Stanford'' and 
		``Log number of trials''. We have also amended the caption to clarify the reason for Stanford being relevant as well as provide a reference to where the predictors are discussed at length in Boyce et al 2023. The revised caption reads:

		\revised{Correlations between an individual predictor and the subjective replication score of the rescue project. The first set of predictors were pre-registered based on the correlates used in Boyce et al. (2023). Original authors at Stanford refers to at the time of the first replication as this may mean the first replication was not independent. For details about how the predictors were coded, see the Methods section of Boyce et al. (2023). The last three predictors were added post-hoc.}

		\theysaid{It is the open science collaborations, not consortium.}

		Oops, thank you for bringing this to our attention, we have fixed it.

@@

		This point is well taken and we have expanded our discussion to acknowledge the larger space of factors at play with replications and rescues and the motivations to do them.

		The concluding part of the discussion now reads:

		\revised{Overall, the diminished and inconsistent effects we found in the rescues suggest that even if a re-replication ``works'', it may be difficult to build upon as follow-up studies will need large samples to detect small effects.
			\newline

			We have focused here on a narrow type of replication and rescue: that done by an early career researcher with the goal of establishing whether this is a result \emph{they} can build on. We acknowledge that this type of replication exists as part of a wider ecosystem of replications that have different properties and are done for different reasons.
			\newline

			In many cases, researchers may have strong prior beliefs about where a replication will succeed, and these priors may influence a researchers interpretation of a failed replication, and thus their calculations about where a re-replication will succeed or is worth attempting. For instance, a researcher who has high credence in the main result is more likely to attribute failure to specific details in the replication, and thus may want to do a rescue where those details are adjusted. On the other hand, a researcher who is dubious of the methods, results, or statistical approaches in an experiment, or who has previously failed to extend the experimental result, may be quicker to interpret a non-replication as a more definitive negative result. What a researcher chooses to do may also depend on the perceived strength of the replication, what size effects they think are relevant, and how important they think the target effect is. Thus, in some cases, when to re-replicate may be a question of prior beliefs and perceived utility rather than just likelihood of a re-replication succeeding.
			\newline

			We opened with a question about what an early-career psychology researcher should do given a failed replication: should they try again or give up and move on?
			In our sample, the odds of a re-replication working are low (consistent with Ebersole et al., 2020).
			Especially if there is not a clear, identifiable candidate cause for replication failure, it may be more efficient to select another result to build on.
		}

@@

		Thank you for this point, we have now clarified that what we meant was had a confidence interval not overlapping 0. The caption now reads

		\revised{Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95\% CI that did not overlap 0; however, in these cases the larger pattern of results was not consistent between original and first replication. }

		\theysaid{So in general, I think this is a neat little paper, which, in my opinion, could benefit from adding a little bit more discussion/depth in the end on what its outcomes could entail.}

@@

		We would like to reiterate our thanks both the editor and the reviewers for their thoughtful and prompt reviews. We hope that our changes to the manuscript and the documentation of our repository offer the necessary clarifications and nuance.
